Extending Morphological Chains with Supervised Learning and Cross-Language Features

### 2015-12-15 paper/paper.tex
butterfly souffle

Let our set of observed words be $D$. We approach this by trying to maximize the likelihood of observing the words in $D$ from the space of all constructible strings from the alphabet, $\Sigma^*$. We maximize the likelihood over $\theta$ as given by

\begin{equation}
    \begin{split}
        L(\theta; D) &= \prod_{w^*\in D} P(w^*) \\
        &= \prod_{w^*\in D} \sum_{z\in C(w^*)} P(w^*, z) \\
        &= \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in\Sigma^*}\sum_{z\in C(w) e^{\theta\cdot\phi(w,z)}}} \\
    \end{split}
\end{equation}

@@

With this substitution, we can formulate the contrastive likelihood as

\begin{equation}
    L_C(\theta; D) = \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}}
\end{equation}

@@

\begin{equation}
    \begin{split}
        L_C(\theta; D) &= \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}} \\
        &\qquad - \lambda \|\theta\|^2
    \end{split}
\end{equation}

Our algorithm operates on the log-likelihood, which takes the following form:

\begin{equation}
    \begin{split}
        L_L(\theta; D) &= \sum_{w^*\in D} \Bigg(\log \sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)} \Bigg.\\
        &\qquad\Bigg. -\log \sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)} \Bigg) \\
        &\qquad - \lambda \|\theta\|^2
    \end{split}
\end{equation}

@@

\begin{equation}
    \begin{split}
        &\frac{\partial L_L(\theta; D)}{\partial \theta_j} \\
        &= \sum_{w^*\in D} \Bigg( \frac{\sum_{z\in C(w^*)} \phi_j(w^*,z)\cdot e^{\theta\cdot\phi(w^*,z)}} {\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} \Bigg. \\
        &\qquad -\frac{\sum_{w\in N(w^*)} \sum_{z\in C(w)} \phi_j(w,z)\cdot e^{\theta\cdot\phi(w,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}} \Bigg) \\
        &\qquad -2\lambda\theta_j
    \end{split}
\end{equation}

We optimize $L_L(\theta; D)$ with its gradient using the LBFGS-B algorithm.

@@

\begin{table*}
    \begin{center}
        \begin{tabular}{ | l | l | l | l | l |}
            \hline
            Language & Method & Precision & Recall & F1 \\ \hline
                     & \citeauthor{Narasimhan15} & 0.807 & 0.722 & 0.762 \\ 
                     & Unsupervised (Baseline) & 0.664  & 0.649 & 0.656 \\ 
            English & Semi-supervised ($\alpha=15$) & 0.892 & 0.762 & 0.822 \\ 
                    & Two language unsupervised & 0.432 & 0.652 & 0.520 \\ 
                    & Two language semi-supervised & 0.449 & 0.759 & 0.564 \\ \hline
                    & \citeauthor{Narasimhan15} & 0.743 & 0.520 & 0.612 \\ 
            Turkish & Unsupervised (Baseline) & 0.675 & 0.371 & 0.479 \\ 
                    & Semi-supervised ($\alpha=50$) & 0.763 & 0.642 & 0.697 \\ \hline
        \end{tabular}
        \caption{Accuracies of Various Models. For our model, we used the top 48000 English words in the wordlist,
        and the MorphoChallenge 2010 Train/Development sets for training and testing, respectively.}
    \end{center}
\end{table*}

@@

\section{Results}
\begin{figure}
    \begin{tikzpicture}
        \begin{axis}[
                height=0.8\linewidth,
                width=\linewidth,
                xmin=1000, xmax=300000,
                ymin=0, ymax=1,
                xmode=log,
                xlabel=Training Vocab Size,
                ylabel=F-1 Score,
            ]
            \addplot coordinates {
                (\engsmallw, 0.556)
                (\engmedw, 0.656)
                (\englargew, 0.383)
            };
            \addlegendentry{English}
            \addplot coordinates {
                (\tursmallw, 0.124)
                (\turmedw, 0.479)
                (\turlargew, 0.520)
            };
            \addlegendentry{Turkish}
        \end{axis}
    \end{tikzpicture}
    \label{fig:unsupervised}
    \caption{Performance of unsupervised model after frequency thresholding training data}
\end{figure}

\begin{figure}
    \begin{tikzpicture}
        \begin{axis}[
                height=0.8\linewidth,
                width=\linewidth,
                xmin=0, xmax=20,
                ymin=0, ymax=1,
                xlabel=$\alpha$,
                ylabel=F-1 Score,
                legend pos=south east,
            ]
            \addplot coordinates {
                (0, 0.656)
                (0.1, 0.683)
                (1, 0.752)
                (3, 0.803)
                (5, 0.815)
                (8, 0.824)
                (10, 0.823)
                (12, 0.821)
                (15, 0.822)
                (20, 0.817)
            };
            \addlegendentry{English}
            \addplot coordinates {
                (0, 0.479)
                (0.1, 0.580)
                (1, 0.663)
                (3, 0.668)
                (5, 0.670)
                (8, 0.680)
                (10, 0.680)
                (20, 0.691)
                (50, 0.697)
            };
            \addlegendentry{Turkish}
        \end{axis}
    \end{tikzpicture}
    \label{fig:supervised}
    \caption{Performance of the semi-supervised model based on $\alpha$.}
\end{figure}

We found that utilizing information from the gold segmentations, we were able to substantially increase our
performance, from an accuracy of 65.5\% to 81.7\% with English, and from 47.9\% to 69.7\% with Turkish.
This is a substantial but expected increase; observing the unsupervised model reveals that it performs
many segmentations with nonexistent affixes, which are rectified by providing input about known affixes.

>>>>

Let our set of observed words be $D$. We approach this by trying to maximize the likelihood of observing the words in $D$ from the space of all constructible strings from the alphabet, $\Sigma^*$. We maximize the log-likelihood over $\theta$ as given by

\begin{equation}
    \begin{split}
        L(\theta; D) &= \log \prod_{w^*\in D} P(w^*) \\
        &= \log \prod_{w^*\in D} \sum_{z\in C(w^*)} P(w^*, z) \\
        &= \log \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in\Sigma^*}\sum_{z\in C(w) e^{\theta\cdot\phi(w,z)}}} \\
    \end{split}
\end{equation}

@@

With this substitution, we can formulate the contrastive log-likelihood as

\begin{equation}
    L_C(\theta; D) = \log \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}}
\end{equation}

@@

\begin{equation}
    \begin{split}
        L_C(\theta; D) &= \log \prod_{w^*\in D} \frac{\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}} \\
        &\qquad - \lambda \|\theta\|^2
    \end{split}
\end{equation}

Simplifying, this becomes:

\begin{equation}
    \begin{split}
        L_C(\theta; D) &= \sum_{w^*\in D} \Bigg(\log \sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)} \Bigg.\\
        &\qquad\Bigg. -\log \sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)} \Bigg) \\
        &\qquad - \lambda \|\theta\|^2
    \end{split}
\end{equation}

@@

\begin{equation}
    \begin{split}
        &\frac{\partial L_C(\theta; D)}{\partial \theta_j} \\
        &= \sum_{w^*\in D} \Bigg( \frac{\sum_{z\in C(w^*)} \phi_j(w^*,z)\cdot e^{\theta\cdot\phi(w^*,z)}} {\sum_{z\in C(w^*)} e^{\theta\cdot\phi(w^*,z)}} \Bigg. \\
        &\qquad -\frac{\sum_{w\in N(w^*)} \sum_{z\in C(w)} \phi_j(w,z)\cdot e^{\theta\cdot\phi(w,z)}} {\sum_{w\in N(w^*)} \sum_{z\in C(w)} e^{\theta\cdot\phi(w,z)}} \Bigg) \\
        &\qquad -2\lambda\theta_j
    \end{split}
\end{equation}

We optimize $L_C(\theta; D)$ with its gradient using the LBFGS-B algorithm.

@@

\begin{table*}
    \begin{center}
        \begin{tabular}{ | l | l | l | l | l |}
            \hline
            Language & Method & Precision & Recall & F1 \\ \hline
                     & \citeauthor{Narasimhan15} & 0.807 & 0.722 & 0.762 \\ 
                     & Unsupervised (Baseline) & 0.744  & 0.756 & 0.750 \\ 
            English & Semi-supervised ($\alpha=15$) & 0.892 & 0.777 & 0.831 \\ 
                    & Two language unsupervised & 0.502 & 0.682 & 0.578 \\ 
                    & Two language semi-supervised & 0.519 & 0.789 & 0.623 \\ \hline
                    & \citeauthor{Narasimhan15} & 0.743 & 0.520 & 0.612 \\ 
            Turkish & Unsupervised (Baseline) & 0.627 & 0.478 & 0.543 \\ 
                    & Semi-supervised ($\alpha=50$) & 0.763 & 0.642 & 0.697 \\ \hline
        \end{tabular}
        \caption{Accuracies of Various Models. For our model, we used the top 48000 English words in the wordlist,
        and the MorphoChallenge 2010 Train/Development sets for training and testing, respectively.}
    \end{center}
\end{table*}

@@

\section{Results}
\begin{figure}
    \begin{tikzpicture}
        \begin{axis}[
                height=0.8\linewidth,
                width=\linewidth,
                xmin=1000, xmax=300000,
                ymin=0, ymax=1,
                xmode=log,
                xlabel=Training Vocab Size,
                ylabel=F-1 Score,
            ]
            \addplot coordinates {
                (\engsmallw, 0.626)
                (\engmedw, 0.746)
                (\englargew, 0.603)
            };
            \addlegendentry{English}
            \addplot coordinates {
                (\tursmallw, 0.124)
                (\turmedw, 0.529)
                (\turlargew, 0.570)
            };
            \addlegendentry{Turkish}
        \end{axis}
    \end{tikzpicture}
    \label{fig:unsupervised}
    \caption{Performance of unsupervised model after frequency thresholding training data}
\end{figure}

\begin{figure}
    \begin{tikzpicture}
        \begin{axis}[
                height=0.8\linewidth,
                width=\linewidth,
                xmin=0, xmax=20,
                ymin=0, ymax=1,
                xlabel=$\alpha$,
                ylabel=F-1 Score,
                legend pos=south east,
            ]
            \addplot coordinates {
                (0, 0.716)
                (0.1, 0.753)
                (1, 0.782)
                (3, 0.803)
                (5, 0.815)
                (8, 0.824)
                (10, 0.823)
                (12, 0.821)
                (15, 0.822)
                (20, 0.817)
            };
            \addlegendentry{English}
            \addplot coordinates {
                (0, 0.549)
                (0.1, 0.620)
                (1, 0.663)
                (3, 0.668)
                (5, 0.670)
                (8, 0.680)
                (10, 0.680)
                (20, 0.691)
                (50, 0.697)
            };
            \addlegendentry{Turkish}
        \end{axis}
    \end{tikzpicture}
    \label{fig:supervised}
    \caption{Performance of the semi-supervised model based on $\alpha$.}
\end{figure}

We found that utilizing information from the gold segmentations, we were able to substantially increase our
performance, from an accuracy of 75\% to 81.7\% with English, and from 54.3\% to 69.7\% with Turkish.
This is a substantial but expected increase; observing the unsupervised model reveals that it performs
many segmentations with nonexistent affixes, which are rectified by providing input about known affixes.
